---
title: "Python Setup in R with Reticulate"
subtitle: "Using R + Python"
author: "Anselmo Araujo"
date: as.character(Sys.Date())
output: 
  html_notebook:
    theme: flatly
    toc: true
    toc_depth: 2
---



Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
knitr::knit_engines$set(
  python = reticulate::eng_python
)
```

```{r}
library(tidyverse)
library(reticulate)
```

```{r}
reticulate::use_virtualenv(virtualenv = file.path('py38'), required = TRUE)
reticulate::py_config()
```

```{python}
import numpy as np
import pandas as pd

np.arange(1, 10)

# Make a sequence in a data frame using dict format
df = pd.DataFrame(data = {"sequence":np.arange(1,20,.01)})

# Use assign (mutate) equivalent to calculate the np.sin() of the series
df = df.assign(value=np.sin(df["sequence"]))

df
```

```{python}
import matplotlib.pyplot as plt

df.plot(x="sequence", y = "value", title = "Matplotlib")

plt.show()
```

```{python}
from sklearn.ensemble import RandomForestClassifier

clf = RandomForestClassifier(random_state=0)

X = [[ 1,  2,  3],  # 2 samples, 3 features
     [11, 12, 13]]

y = [0, 1]  # classes of each sample

clf.fit(X, y)
```

```{python}
clf.predict(X) 
```

```{python}
from sklearn.cluster import AffinityPropagation
from sklearn.datasets import make_blobs

# #############################################################################
# Generate sample data
centers = [[1, 1], [-1, -1], [1, -1]]
X, labels_true = make_blobs(n_samples=300, centers=centers, cluster_std=0.5,
                            random_state=0)

# Compute Affinity Propagation
af = AffinityPropagation(preference=-50).fit(X)
cluster_centers_indices = af.cluster_centers_indices_
labels = af.labels_

n_clusters_ = len(cluster_centers_indices)

# #############################################################################
# Plot result
import matplotlib.pyplot as plt
from itertools import cycle

plt.close('all')
plt.figure(1)
plt.clf()

colors = cycle('bgrcmykbgrcmykbgrcmykbgrcmyk')
for k, col in zip(range(n_clusters_), colors):
    class_members = labels == k
    cluster_center = X[cluster_centers_indices[k]]
    plt.plot(X[class_members, 0], X[class_members, 1], col + '.')
    plt.plot(cluster_center[0], cluster_center[1], 'o', markerfacecolor=col,
             markeredgecolor='k', markersize=14)
    for x in X[class_members]:
        plt.plot([cluster_center[0], x[0]], [cluster_center[1], x[1]], col)

plt.title('Estimated number of clusters: %d' % n_clusters_)
plt.show()
```

```{python}
# TensorFlow and tf.keras
import tensorflow as tf
from tensorflow import keras

# Helper Libraries
import numpy as np
import matplotlib.pyplot as plt

print('Tensorflow version ' + tf.__version__)

# Load the fashion_mnist dataset from keras
fashion_mnist = keras.datasets.fashion_mnist
(train_images, train_labels), (test_images, test_labels) = fashion_mnist.load_data()

train_images.shape

np.unique(train_labels)

class_names = ['T-shirt/top', 'Trouser', 'Pullover', 'Dress', 'Coat',
               'Sandal', 'Shirt', 'Sneaker', 'Bag', 'Ankle boot']
               
plt.figure()
plt.imshow(train_images[1])
plt.colorbar()
plt.grid(False)
plt.show()
```

```{python}
plt.figure(figsize=(10,10))
for i in range(25):
    plt.subplot(5,5,i+1)
    plt.xticks([])
    plt.yticks([])
    plt.grid(False)
    plt.imshow(train_images[i], cmap=plt.cm.binary)
    plt.xlabel(class_names[train_labels[i]])
plt.show()
```

#### Modeling with Keras

Make a *keras* model using the *Sequential()* with 3 steps: Flatten, Dense and Dense.

```{python}
model = keras.Sequential([
    keras.layers.Flatten(input_shape=(28, 28)),
    keras.layers.Dense(128, activation='relu'),
    keras.layers.Dense(10)
])

# compile the model with the 'adam' optimizer
model.compile(
    optimizer = 'adam',
    loss      = tf.keras.losses.SparseCategoricalCrossentropy(from_logits=True),
    metrics   = ['accuracy']
)

# Inspect the model summary
model.summary()
```

Fit the Keras Model
**Critical Step**

```{python}
model.fit(train_images, train_labels, epochs=10, verbose=1)
```

Training History

```{python}
history = model.history.history
history
```
Plot history

```{r}
py$history %>% 
    as_tibble() %>%
    unnest(loss, accuracy) %>%
    rowid_to_column() %>%
    pivot_longer(-rowid) %>%
    ggplot(aes(rowid, value, color = name)) +
    geom_line() +
    geom_point() +
    labs(title = "Training Accuracy")
```

Evaluate accuracy on the out-of-sample images

```{python}
test_loss, test_acc = model.evaluate(test_images,  test_labels, verbose=2)
```

Make Predictions

The model produces linear outputs cakked “logits”. The softmax layer to converts the logits to probabilities.

```{python}
probability_model = tf.keras.Sequential([model, tf.keras.layers.Softmax()])
```

We can then classify all of the test images (held out)

```{python}
predictions = probability_model.predict(test_images)
```

We can make a prediction for the first image.

```{python}
predictions[0]
```

Use np.argmax() to determine which index has the highest probability.

```{python}
np.argmax(predictions[0])
```

The index value can be retrieved with np.max().

```{python}
np.max(predictions[0])
```

Get the class name.

```{python}
class_names[np.argmax(predictions[0])]
```

And visualize the image.

```{python}
plt.figure()
plt.imshow(test_images[0])
plt.colorbar()
plt.grid(False)
plt.show()
```

### Scraping ETF Constituents with Python from RStudio
[link](https://robotwealth.com/using-python-from-the-comfort-of-r-studio/)

```{python}
import pandas as pd
def get_holdings(spdr_ticker):
    url = f'http://www.sectorspdr.com/sectorspdr/IDCO.Client.Spdrs.Holdings/Export/ExportCsv?symbol={spdr_ticker}'
    df = pd.read_csv(url, skiprows=1).to_csv(f'{spdr_ticker}_holdings.csv', index=False)
    return df
```

```{python}
import pandas as pd
def get_holdings(spdr_ticker):
    """read in ETF holdings"""
    url = f"http://www.sectorspdr.com/sectorspdr/IDCO.Client.Spdrs.Holdings/Export/ExportCsv?symbol={spdr_ticker}"
    df = pd.read_csv(url, skiprows=1, usecols=[i for i in range(3)]) 
    return df
```

